Supplemental Information A: Full Annotated Analysis and Reproducible Code

### 2022-06-20 Supplemental-Material-A-Annotated-Analysis.Rmd
sample selection v.02

rownames(dtS1Availability) <- paste("PP", 1:nrow(dtS1Availability), sep = "_")
colnames(dtS1Availability) <- paste("t", 1:ncol(dtS1Availability), sep = "_")
dtS1Availability[1:ncol(dtS1Availability)] <- lapply(dtS1Availability[1:ncol(dtS1Availability)], function(x) {
    cell_spec(x,
              bold = FALSE,
              color = "white",
              background = ifelse(x == 1, "green", "red")
              )
})
kbl(
  dtS1Availability,
  format = "html",
  escape = FALSE,
  align = "c",
  booktabs = TRUE,
  caption = "Study 1: Data Availability" # complete caption for main document
) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")

@@

```{r testOptimizer, include=FALSE}
set.seed(42)
df <- as.data.table(matrix(rbinom(5*10,1,.66), nrow=5, ncol=10))
df

best.list.row.df = list()
for (i in 1:nrow(df)) {
  # get best subset for rows based on how many columns have more than 66% data
  rowlist = combn(nrow(df), i)
  numobs = apply(rowlist, 2, function(x) sum(colMeans(df[x,])*100 >= 66))
  cat("For subsets of", i, "rows, the highest number of observations is", max(numobs), "out of the", ncol(df), "maximum. Product =", i*max(numobs),"\n")
  best = which(numobs == max(numobs))[1]
  best.list.row.df = c(best.list.row.df, list(rowlist[, best]))
}

best.list.col.df = list()
for (i in 1:ncol(df)) {
  # get best subset for columns based on how many rows have more than 66% data
  collist = combn(ncol(df), i)
  numobs = apply(collist, 2, function(x) sum(rowMeans(df[, ..x])*100 >= 66))
  cat("For subsets of", i, "columns, the highest number of participants is", max(numobs), "out of the", nrow(df), "maximum. Product =", i*max(numobs),"\n")
  best = which(numobs == max(numobs))[1]
  best.list.col.df = c(best.list.col.df, list(collist[, best]))
}
rm(df, best.list.row.df, best.list.col.df)
```

@@

dtS2Availability[1:ncol(dtS2Availability)] <- lapply(dtS2Availability[1:ncol(dtS2Availability)], function(x) {
    cell_spec(x,
              bold = FALSE,
              color = "white",
              background = ifelse(x == 1, "green", "red")
              )
})
kbl(
  dtS2Availability,
  format = "html",
  escape = FALSE,
  align = "c",
  booktabs = TRUE,
  caption = "Study 2: Data Availability" # complete caption for main document
) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")
```

@@

dtS3Availability[1:ncol(dtS3Availability)] <- lapply(dtS3Availability[1:ncol(dtS3Availability)], function(x) {
    cell_spec(x,
              bold = FALSE,
              color = "white",
              background = ifelse(x == 1, "green", "red")
              )
})
kbl(
  dtS3Availability,
  format = "html",
  escape = FALSE,
  align = "c",
  booktabs = TRUE,
  caption = "Study 2: Data Availability" # complete caption for main document
) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")
```

@@

varNamesDaily <- varNames %>%
  filter(
    surveyS1 == "daily",
    surveyS2 == "daily",
    surveyS3 == "daily"
  ) %>%
  select(varNam)
varNamesDailyFull <-
  data.frame(varNam = Reduce(intersect, list(
    names(dtS1$full), names(dtS3$full), names(dtS3$full)
  ))) %>%
  filter(!str_detect(varNam, ".pre$|.post$"))

varNamesPost <- varNames %>%
  filter(
    surveyS1 == "post",
    surveyS2 == "post",
    surveyS3 == "post"
  ) %>%
  select(varNam)
varNamesPostFull <-
  data.frame(varNam = Reduce(intersect, list(
    names(dtS1$full), names(dtS3$full), names(dtS3$full)
  ))) %>%
  filter(str_detect(varNam, ".post$"))
```

@@

dtMain3mpca <- rbind(
  dtS1$full %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S1"),
  dtS3$full %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S2"),
  dtS3$full %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S3")
  ) %>%
  group_by(study, PID) %>%
  mutate(ID = cur_group_id()) %>%
  ungroup %>%
  mutate(
    date = as.Date(gsub(" .*", "", TID)),
    week = strftime(date, format = "%Y-W%V")
    ) %>%
  select(
    ID,
    PID,
    TID,
    date,
    week,
    TIDnum,
    study,
    everything()
  ) %>%
  arrange(ID, TIDnum)

@@

dtMain3mpcaWeekly <- dtMain3mpca %>%
  group_by(ID, week, study) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
  ungroup %>%
  group_by(study) %>%
  mutate(TIDnum = as.numeric(factor(week))) %>%
  ungroup %>%
  select(ID, week, TIDnum, everything()) %>%
  mutate_all(~ifelse(is.nan(.), NA, .))

dtMain3mpcaWeekly

stats::reshape(dtMain3mpca, idvar = "ID", timevar = "TIDnum", direction = "wide")

@@

>>>>

rownames(dtS1Availability) <- paste("PP", 1:nrow(dtS1Availability), sep = "_")
colnames(dtS1Availability) <- paste("t", 1:ncol(dtS1Availability), sep = "_")
dtS1AvailabilityKbl <- dtS1Availability

dtS1AvailabilityKbl[1:ncol(dtS1AvailabilityKbl)] <- lapply(dtS1AvailabilityKbl[1:ncol(dtS1AvailabilityKbl)], function(x) {
    cell_spec(x,
              bold = FALSE,
              color = "white",
              background = ifelse(x == 1, "green", "red")
              )
})
kbl(
  dtS1AvailabilityKbl,
  format = "html",
  escape = FALSE,
  align = "c",
  booktabs = TRUE,
  caption = "Study 1: Data Availability" # complete caption for main document
) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")

@@

```{r testOptimizer, include=FALSE}
# set.seed(42)
# nr <- 5
# nc <- 10
# df <- as.data.table(matrix(rbinom(nr*nc,1,.66), nrow=nr, ncol=nc))
# df
# 
# best.list.row.df = list()
# for (i in 1:nrow(df)) {
#   # get best subset for rows based on how many columns have more than 66% data
#   rowlist = combn(nrow(df), i)
#   numobs = apply(rowlist, 2, function(x) sum(colMeans(df[x,])*100 >= 66))
#   cat("For subsets of", i, "rows, the highest number of observations is", max(numobs), "out of the", ncol(df), "maximum. Product =", i*max(numobs),"\n")
#   best = which(numobs == max(numobs))[1]
#   best.list.row.df = c(best.list.row.df, list(rowlist[, best]))
# }
# 
# best.list.col.df = list()
# for (i in 1:ncol(df)) {
#   # get best subset for columns based on how many rows have more than 66% data
#   collist = combn(ncol(df), i)
#   numobs = apply(collist, 2, function(x) sum(rowMeans(df[, ..x])*100 >= 66))
#   cat("For subsets of", i, "columns, the highest number of participants is", max(numobs), "out of the", nrow(df), "maximum. Product =", i*max(numobs),"\n")
#   best = which(numobs == max(numobs))[1]
#   best.list.col.df = c(best.list.col.df, list(collist[, best]))
# }
# rm(df, best.list.row.df, best.list.col.df)
```

```{r testOptimizer03, include=FALSE}
# library(dplyr)
# library(ROI)
# library(ROI.plugin.glpk)
# library(ompr)
# library(ompr.roi)
# 
# set.seed(42)
# tf <- matrix(sample(c(TRUE, FALSE), 1488, replace=TRUE), 31)
# M3 <- t(replicate(31, 1969:2016, simplify=TRUE))
# M3[tf] <- NA
# 
# m <- +!is.na(M3) # gets logical matrix; 0 if NA else 1    
# #m <- as.data.frame(dtS1AvailabilityTest)
# nr <- nrow(m)
# nc <- ncol(m)
# n_years <- 15 
# 
# 
# model <- MIPModel() %>% 
#   # keep[i,j] is 1 if matrix cell [i,j] is to be kept else 0
#   add_variable(keep[i,j], i = 1:nr, j = 1:nc, typ = "binary") %>% 
#   # rm_row[i] is 1 if row i is selected for removal else 0
#   add_variable(rm_row[i], i = 1:nr, type = "binary") %>% 
#   # rm_col[j] is 1 if column j is selected for removal else 0
#   add_variable(rm_col[j], j = 1:nc, type = "binary") %>% 
#   # maximize good cells kept
#   set_objective(sum_expr(keep[i,j], i = 1:nr, j = 1:nc), "max") %>% 
#   # cell can be kept only when row is not selected for removal
#   add_constraint(sum_expr(keep[i,j], j = 1:nc) <= 1 - rm_row[i], i = 1:nr) %>%
#   # cell can be kept only when column is not selected for removal
#   add_constraint(sum_expr(keep[i,j], i = 1:nr) <= 1 - rm_col[j], j = 1:nc) %>%
#   # keep at most n_years columns i.e. remove at least (nc - n_years) columns
#   # only non-NA values can be kept
#   add_constraint(m[i,j] + rm_row[i] + rm_col[j] >= 1, i = 1:nr, j = 1:nc) %>% 
#   #add_constraint(sum_expr(m[,j], j = 1:nc)/nc*100 >= 66) %>% 
#   #add_constraint(sum_expr(m[i,], i = 1:nr)/nr*100 >= 66) %>% 
#   # keep cols with more 
#   #add_constraint(colMeans(m[i,j], i = 1:nr, j = 1:nc) >= 0.66) %>%
#   # I used >= instead of == to avoid infeasiblity
#   #add_constraint(sum_expr(rm_col[j], j = 1:nc) >= nc - n_years) %>% 
#   # solve using free glpk solver
#   solve_model(with_ROI(solver = "glpk", verbose = TRUE))
# 
# solver_status(model)
# rm_rows <- model %>% 
#   get_solution(rm_row[i]) %>% 
#   filter(value > 0) %>% 
#   pull(i)
# 
# rm_cols <- model %>% 
#   get_solution(rm_col[j]) %>% 
#   filter(value > 0) %>% 
#   pull(j) 
# 
# result <- m[-rm_rows, -rm_cols, drop = F]
# result
```

```{r testOptimizer05}
cleanM <- function(M, c = 66) {
  check <- list()
  rowNamesOut <- c()
  colNamesOut <- c()
  for (i in 1:length(c(row.names(M), names(M)))) {
    rowMeans <- rowMeans(M) * 100
    colMeans <- colMeans(M) * 100
    rcMeans <- c(rowMeans, colMeans)
    rm <- which.min(rcMeans) %>% names
    rc <- ifelse(startsWith(rm, "PP_"), "row", "col")

    check[[i]] <- rcMeans

    if (!all(rcMeans >= c) && rc == "row") {
      M <- M[!(row.names(M) %in% rm), ]
      rowNamesOut <- append(rowNamesOut, rm)
      cat(i, ": Row ", rm, " had a completion rate of ", format(round(min(rcMeans), 2), nsmall = 2), "% and was removed.\n", sep = "")
    } else if (!all(rcMeans >= c) && rc == "col") {
      M <- M[, !(names(M) %in% rm)]
      colNamesOut <- append(colNamesOut, rm)
      cat(i, ": Column ", rm, " had a completion rate of ", format(round(min(rcMeans), 2), nsmall = 2), "% and was removed.\n", sep = "")
    } else {
      cat(i, ": All row- and column means are over ", c, "%. The final matrix has ", nrow(M), " rows and ", ncol(M), " columns.", sep = "")
      return(
        list(
          reducedMatrix = M,
          rowNamesIn = row.names(M),
          colNamesIn = names(M),
          rowNamesOut = rowNamesOut,
          colNamesOut = colNamesOut,
          rowMeans = rowMeans,
          colMeans = colMeans
        )
      )
    }
  }
}

dtS1RedInfo <- cleanM(M = as.data.frame(dtS1Availability), c = 66)

PIDout <- gsub("PP_", "", dtS1RedInfo$rowNamesOut) %>% as.numeric
TIDout <- gsub("t_", "", dtS1RedInfo$colNamesOut) %>% as.numeric

dtS1Red <- dtS1$full %>%
  filter(
    !PID %in% PIDout,
    !TIDnum %in% TIDout
  ) %>%
  mutate(TIDnum = as.numeric(factor(TIDnum)))
rm(PIDout, TIDout)

dtS1AvailabilityRedKbl <- dtS1RedInfo$reducedMatrix
dtS1AvailabilityRedKbl[1:ncol(dtS1AvailabilityRedKbl)] <- lapply(dtS1AvailabilityRedKbl[1:ncol(dtS1AvailabilityRedKbl)], function(x) {
    cell_spec(x,
              bold = FALSE,
              color = "white",
              background = ifelse(x == 1, "green", "red")
              )
})
kbl(
  dtS1AvailabilityRedKbl,
  format = "html",
  escape = FALSE,
  align = "c",
  booktabs = TRUE,
  caption = "Study 1: FinalData Availability" # complete caption for main document
) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")
```

@@

dtS2AvailabilityKbl <- dtS2Availability
dtS2AvailabilityKbl[1:ncol(dtS2AvailabilityKbl)] <- lapply(dtS2AvailabilityKbl[1:ncol(dtS2AvailabilityKbl)], function(x) {
    cell_spec(x,
              bold = FALSE,
              color = "white",
              background = ifelse(x == 1, "green", "red")
              )
})
kbl(
  dtS2AvailabilityKbl,
  format = "html",
  escape = FALSE,
  align = "c",
  booktabs = TRUE,
  caption = "Study 2: Data Availability" # complete caption for main document
) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")
```

```{r s2SampleReduction}
dtS2RedInfo <- cleanM(M = as.data.frame(dtS2Availability), c = 66)

PIDout <- gsub("PP_", "", dtS2RedInfo$rowNamesOut) %>% as.numeric
TIDout <- gsub("t_", "", dtS2RedInfo$colNamesOut) %>% as.numeric

dtS2Red <- dtS2$full %>%
  filter(
    !PID %in% PIDout,
    !TIDnum %in% TIDout
  ) %>%
  mutate(TIDnum = as.numeric(factor(TIDnum)))
rm(PIDout, TIDout)

dtS2AvailabilityRedKbl <- dtS2RedInfo$reducedMatrix
dtS2AvailabilityRedKbl[1:ncol(dtS2AvailabilityRedKbl)] <- lapply(dtS2AvailabilityRedKbl[1:ncol(dtS2AvailabilityRedKbl)], function(x) {
    cell_spec(x,
              bold = FALSE,
              color = "white",
              background = ifelse(x == 1, "green", "red")
              )
})
kbl(
  dtS2AvailabilityRedKbl,
  format = "html",
  escape = FALSE,
  align = "c",
  booktabs = TRUE,
  caption = "Study 2: Final Data Availability" # complete caption for main document
) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")
```

@@

dtS3AvailabilityKbl <- dtS3Availability
dtS3AvailabilityKbl[1:ncol(dtS3AvailabilityKbl)] <- lapply(dtS3AvailabilityKbl[1:ncol(dtS3AvailabilityKbl)], function(x) {
    cell_spec(x,
              bold = FALSE,
              color = "white",
              background = ifelse(x == 1, "green", "red")
              )
})
kbl(
  dtS3AvailabilityKbl,
  format = "html",
  escape = FALSE,
  align = "c",
  booktabs = TRUE,
  caption = "Study 2: Data Availability" # complete caption for main document
) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")
rm(dtS3AvailabilityKbl)
```

```{r s3SampleReduction}
dtS3RedInfo <- cleanM(M = as.data.frame(dtS3Availability), c = 66)

PIDout <- gsub("PP_", "", dtS3RedInfo$rowNamesOut) %>% as.numeric
TIDout <- gsub("t_", "", dtS3RedInfo$colNamesOut) %>% as.numeric

dtS3Red <- dtS3$full %>%
  filter(
    !PID %in% PIDout,
    !TIDnum %in% TIDout
  ) %>%
  mutate(TIDnum = as.numeric(factor(TIDnum)))
rm(PIDout, TIDout)

dtS3AvailabilityRedKbl <- dtS3RedInfo$reducedMatrix
dtS3AvailabilityRedKbl[1:ncol(dtS3AvailabilityRedKbl)] <- lapply(dtS3AvailabilityRedKbl[1:ncol(dtS3AvailabilityRedKbl)], function(x) {
    cell_spec(x,
              bold = FALSE,
              color = "white",
              background = ifelse(x == 1, "green", "red")
              )
})
kbl(
  dtS3AvailabilityRedKbl,
  format = "html",
  escape = FALSE,
  align = "c",
  booktabs = TRUE,
  caption = "Study 3: Final Data Availability" # complete caption for main document
) %>%
  kable_classic(
    full_width = FALSE,
    lightable_options = "hover",
    html_font = "Cambria"
  ) %>%
  scroll_box(width = "100%", height = "500px")
```

@@

varNamesDaily <- varNames %>%
  filter(
    surveyS1 == "daily",
    surveyS2 == "daily",
    surveyS3 == "daily"
  ) %>%
  select(varNam)
varNamesDailyFull <-
  data.frame(varNam = Reduce(intersect, list(
    names(dtS1Red), names(dtS3Red), names(dtS3Red)
  ))) %>%
  filter(!str_detect(varNam, ".pre$|.post$"))

varNamesPost <- varNames %>%
  filter(
    surveyS1 == "post",
    surveyS2 == "post",
    surveyS3 == "post"
  ) %>%
  select(varNam)
varNamesPostFull <-
  data.frame(varNam = Reduce(intersect, list(
    names(dtS1Red), names(dtS3Red), names(dtS3Red)
  ))) %>%
  filter(str_detect(varNam, ".post$"))
```

@@

dtMain3mpca <- rbind(
  dtS1Red %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S1"),
  dtS3Red %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S2"),
  dtS3Red %>% select(any_of(varNam3mpcaMain)) %>% mutate(study = "S3")
  ) %>%
  group_by(study, PID) %>%
  mutate(ID = cur_group_id()) %>%
  ungroup %>%
  mutate(
    date = as.Date(gsub(" .*", "", TID)),
    week = strftime(date, format = "%Y-W%V")
    ) %>%
  select(
    ID,
    PID,
    TID,
    date,
    week,
    TIDnum,
    study,
    everything()
  ) %>%
  arrange(ID, TIDnum)

@@

dtMain3mpcaWeekly <- dtMain3mpca %>%
  group_by(ID, week, study) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE) %>%
  ungroup %>%
  group_by(study) %>%
  mutate(TIDnum = as.numeric(factor(week))) %>%
  ungroup %>%
  select(ID, week, TIDnum, everything()) %>%
  mutate_all(~ifelse(is.nan(.), NA, .))
```

### Timescale ML test

```{r }
dtMain3mpca <- dtMain3mpca %>%
  mutate(
    TIDdate = as.numeric(factor(date)),
    TIDweek = as.numeric(factor(week))
  )

test <-
  lme4::lmer(KeyNeedFulfillment ~ 1 + (1 + TIDnum + TIDdate + TIDweek | PID),
    data = dtMain3mpca
  ) # use optim if it does not converge

test2 <-
  lme(
    KeyNeedFulfillment ~ 1,
    random = ~ 1 + TIDnum + TIDdate + TIDweek | PID,
    data = dtMain3mpca %>% filter(!is.na(KeyNeedFulfillment)),
    control = list(opt = "nlmimb")
  ) # use optim if it does not converge

summary(test2)
#summ(test, digits = 3, center = FALSE)
anova(test2)

@@

## Multiple Imputation

```{}

## Center and Normalize

```{r mainCentered}
varNam3mpcaMainAnalysis <- varNamesDailyFull %>%
  filter(!varNam %in% c(dropVarPcaMain, "PID", "TID", "TIDnum")) %>%
  pull

dtMain3mpcaNorm <-
  PCAnorm(data = dtMain3mpca,
          pid = "ID",
          tid = "TIDnum",
          selection = varNam3mpcaMainAnalysis)
```

## Long to Wide

```{r longToWide}

dtMain3mpcaWideGmc <- dtMain3mpcaNorm %>% 
  select(
    ID,
    TIDnum,
    ends_with("_gmc")
  ) %>%
  reshape2::melt(.,
              id.vars = c("ID", "TIDnum")) %>%
  mutate(variable = paste(variable, TIDnum, sep = "_")) %>%
  select(-TIDnum) %>% 
  arrange(variable, ID) %>%
  reshape(., idvar = "ID", timevar = "variable", direction = "wide")
  #pivot_wider(names_from = variable, values_from = value)

## Three-Way ANOVA

```{r threeWayAnova}
library(ThreeWay)
# threewayanova(
#   Y = dtMain3mpcaWideGmc,
#   n = nrow(dtMain3mpcaWideGmc), # Nr. ppt.s
#   m = length(unique(dtMain3mpcaNorm$TIDnum)), # Nr. time points
#   p = length(varNam3mpcaMainAnalysis) # Nr. variables
# )
```

## Plot (x = var, y = z, group/line = t)
